Compararison of analytical and artificial neural network models in predicting hot deformation behavior of AISI P20 and interpolation--extrapolation capacity validation

### 2022-10-12 Article_PTM.tex
updated version of .tex file

\title{Phenomelogical and artificial neural network models' comparison study for hot deformation of AISI P20: Interpolation and extrapolation predictions}

\author[LGP]{Pierre Tize Mha}
\author[LGP]{Amèvi Tongne}
\author[LGP]{Olivier Pantalé \corref{cor1}}
\ead{Olivier.Pantale@enit.fr}
\ead[url]{http://www.enit.fr}

@@

\address[LGP]{Laboratoire Génie de Production, INP/ENIT, Université de Toulouse, 47 Av d'Azereix, Tarbes, France 65016}

\begin{abstract}
Non-linear numerical modeling is increasingly used for the simulation of complex processes such as forming or machining. The constitutive laws necessary for these simulations are therefore becoming more and more complex regarding the increasingly precise consideration of physical phenomena (plasticity, thermal dependence, damage ...) and their implementation in numerical codes requires the identification of many parameters. The use of these constitutive laws for dynamic or quasi-static applications requires the identification of these parameters under conditions close to those encountered during the real process, mainly in terms of deformation, strain rates, and temperatures. The availability of a constitutive law in the finite element codes does not guarantee the success of a study to be carried out if it does not comply with the model development criteria. This is because the models integrated in the numerical codes are not developed based on forms that are adjustable to any type of study. In this study, a new form of non-linear constitutive flow law based on the Modified Zerilli--Armstrong model, which can answer the above problem, has been developed to apply it to the numerical simulation of two different tests (a quasi-static compression test, the necking of a circular bar). This flow law is based on the modified Zerilli--Armstrong model, which, together with the new modified Johnson--Cook model, has been compared to appreciate the relevance of the proposal. For that, an implementation of this new law via the VUHARD subroutine into the Abaqus/Explicit finite element code was made to model the two tests. The comparison of the results obtained (from identification) by our proposed law with those obtained using the NMJC shows that this new law better approaches the experiments than the other one. This is also showed through the numerical results using the Abaqus software. It can be said that this way of formulating a flow law allows to highlight the great performance of the proposed approach. Although this law has been used only used for quasi-static tests, we can say that it can also be used in dynamic tests.
\end{abstract}

\begin{keyword}
Zerilli--Armstrong flow law \sep Constitutive Behavior \sep Finite Element Method \sep Numerical Implementation \sep Johnson--Cook flow law
\end{keyword}

@@

%--------------------------------------------------------------------------
\section{Introduction\label{sec:Introduction}}
%--------------------------------------------------------------------------
\textcolor{red}{Due to technological advances, industries are constantly looking for innovations to maintain their competitiveness. For the metallurgical industry, these innovations involve the development of new materials. In conjunction with the implementation of these new materials, it becomes necessary to characterize their thermomechanical behavior, in particular in order to feed the data bases used by the numerical simulation models more and more widespread nowadays \cite{Trimble-2020-Finite, Yuan-2021-Particle}}.

\textcolor{red}{Among the different approaches used, the finite element method (FEM) is one of the numerical methods that has become the most popular nowadays. Like any numerical method, the accuracy of the solution depends very strongly on the constitutive equation describing the behavior of the material and implemented in the computational code \cite{Li-2021-Finite, Liu-2021-Finite}. This constitutive equation is linked to the mathematical model that describes it. This choice is not entirely free, insofar as the choice of the behavior law (speaking of an analytically formulated model) used to describe the behavior of a material is guided by its availability in the finite element codes \cite{Miller-2021-Thermomechanical}. This forces an adaptation between the studied material (which is real) and the behavior law (representative mathematical model) available in the FEM code which is not necessarily the one able to describe the behavior of the studied material. The problem of the non flexibility of the models available in the finite element codes appears immediately and limits the adequacy of the model to the reality of the behavior since the behavior laws established and available in the FEM codes are for the most part specific to a particular material or class of materials. In other words, as soon as they are applied to materials different from the one for which they were originally developed, they find their limits because they do not manage to represent the reality described by the experiment as one would wish. It is therefore interesting to think about developing a method that can be applied to any material of any class and implemented in a finite element code. This approach involves a number of parameters depending on the material studied, but with a more global mathematical form adaptable to a wider range of materials. In this study, we propose an approach allowing a better adaptability of the behavior law to various materials. To justify the relevance of this proposal, let us first examine the models developed so far and try to see if some of them meet this need}.

\textcolor{red}{According to the literature, analytical models describing the behavior of materials can be grouped into three categories: empirical models, semi-empirical models, and physical models. Empirical models describe the behavior of the material by considering macroscopic quantities without considering its microstructure, whereas physical models consider the microstructure. As for the semi-empirical models, they try to be at the border between the empirical and physical models. Among a large number of models, the Johnson--Cook (JC) \cite{Johnson-1983-Constitutive, Johnson-1985-Fracture} and the Zerilli--Armstrong (ZA) \cite{Zerilli-1987-Dislocation} models are the best known and available in finite element codes. The JC model is widely used because it is simple to use and has fewer constants to be determined \cite{Khan-2004-Quasi-static, NematNasser-2003-Thermomechanical}. However, this model, in its original state, contains some problems related to the phenomena describing the behavior of the material. We refer here to strain hardening, strain rate hardening, and thermal effects that are described separately in the JC model as analyzed by Jia \eal \cite{Jia-2021-Modified-JC}. To overcome this difficulty, several modified forms of the Johnson--Cook model have been developed in the past \cite{Rule-1998-Revised-JC, Vural-2003-Large, Lin-2010-Modified-JC, Lin-2012-Phenomenological, Li-2013-Modified-JC, Zhou-2019-Research, Zhang-2015-Modified-JC}}.

\textcolor{red}{Due to its more physical formulation, the ZA model is preferred over the JC model because it allows for the coupling between strain rate and temperature effects \cite{Johnson-1988-Evaluation, Voyiadjis-2005-Microstructural, Dey-2007-Influence}. However, the drawbacks of this model are related to the temperature limit value of the test. Indeed, this model applies only to low strain rates and for temperatures limited to $0.6~T_m$ (where $T_m$ is the melting temperature of the material) \cite{Chiou-2005-Strain, Lee-2005-Strain, Chen-2007-Comparative, Lee-2006-Effects}. Moreover, this model, in its original form, does not take into account the coupling between deformation and temperature effects \cite{Samantaray-2009-Thermo-viscoplastic}. In order to reduce these shortcomings (\eg applying this model over a wider temperature range) of the ZA model, Modified Zerilli--Armstrong forms (MZA) have been proposed by several authors over the last past years \cite{Nemat-2004-Plasticity, Lennon-2004-Influence, Muralli-2017-Performance, Cheng-2021-Modified-ZA, Muralli-2021-Finite}}.

\textcolor{red}{These two approaches are among the most popular and the choice depends on the effectiveness of the method for a given problem. So we have on the one hand, the JC formulation and its variants and on the other hand the ZA approach with its variants too. Another approach proposed by Lin \eal \cite{Lin-2010-Modified-JC} is to combine the JC model with the MZA model (in its modified form proposed by Samantaray \eal \cite{Samantaray-2009-Thermo-viscoplastic}). This approach allows combining on the one hand the factor taking into account the strain hardening effects described by the JC model and on the other hand the factor describing the strain rate effects and the thermal softening described by the MZA model in a unified formulation. The major difficulty of all the variants mentioned above is the restriction of the parameters on which the model depends, thus limiting the applicability domain of the developed model. Therefore, it seems necessary to look for a method that does not limit the number of parameters, unlike the approaches mentioned above, and that can describe more accurately the reality observed by the experiment}.

\textcolor{red}{Artificial neural networks (ANNs) offer an alternative way to solve these problems. ANNs are currently used to solve problems that are difficult to conceptualize using traditional computational methods. Thus, unlike a classical approach based on a regression method, an artificial neural network does not need to know the mathematical form of the model it seeks to reproduce and can therefore learn directly from experimental data without any prior assumption on the nature of the input parameters and their interactions. Thus, artificial neural networks can enable new approaches for modeling material behavior and have been successfully applied for the prediction of constitutive relationships of some metals and alloys \cite{Lin-2008-ANN, Lu-2011-ANN, Ashtiani-2016-CSP, Huang-2021-CMM}. Although this approach is very effective, it does not allow the interpretation of the obtained results, because it does not take into account the physics of the phenomena for the considered material}.

\textcolor{red}{From the above, we can retain that there are two main families of models to describe the behavior of a material. On the one hand, analytical models presenting a problem of flexibility and on the other hand, implicit models, based on neural networks, more adaptable but presenting a problem of physical interpretation. It is therefore interesting to develop a physical model that is able to adapt to any material and can be efficiently implemented in a finite element code. In this perspective, we propose here a new formulation based on the MZA model, whose interest is to answer the two problems mentioned above. To appreciate the relevance of our approach, we used the Abaqus/Explicit software which offers the possibility to implement a new behavior law using a FORTRAN VUMAT or VUHARD subroutine following the example of the works proposed by \cite{Gao-2007-FRT, JansenVanRensburg-2012-TSV, Duc-Toan-2012-MJC, Ming-2018-ERV}}.

\textcolor{red}{The main paragraphs of this work are as follows: Section \ref{sec:ConstLaws} presents two  known constitutive laws which will serve as the basis of the the proposed one presented in Section \ref{sec:ProposedModel}. Since this model is new and will be used to perform a numerical model using Abaqus, the three derivatives of flow stress $\sigma$ with respect to the strain $\varepsilon^p$, the strain rate $\mdot{\varepsilon}^p$ and temperature are required. So, determination of these derivatives is done in this Section. Section \ref{sec:Application} is devoted to the comparison of the proposed model to the others one where identification is done on a 100Cr6 steel. Section \ref{sec:NumSimulations} details the numerical results of the models identified in Section \ref{sec:Application} using the cylindrical compression test and the necking of circular benchmark tests. Conclusion and future works are detailed in Section \ref{sec:Conclusion}}.

%----------------------------------------------
\section{Experimental procedure}
%----------------------------------------------
The material used in this study is the modified carbon alloy AISI P20 whose chemical composition is given in the Table. Cylindrical specimens were machined with an initial diameter of $10$ mm and a height of $15$ mm. Hot compression tests were conducted on a Gleeble-3800 thermo-simulator (Figure \ref{fig:Gleeble3800}), at temperatures of $1050^\circ$C,  $1050^\circ$C,  $1100^\circ$C,  $1150^\circ$C,  $1200^\circ$C and  $1250^\circ$C with strain rates of $0.001\ \text{s}^{-1}$, $0.01\ \text{s}^{-1}$, $0.1\ \text{s}^{-1}$, $1.0\ \text{s}^{-1}$, $2.0\ \text{s}^{-1}$ and $5.0\ \text{s}^{-1}$. Thin tantalum sheets were used as a lubricant material at the contact surface of anvils and specimen to minimize the frictions during hot deformation. The samples were heated to $1260^\circ$C temperature and were held for $5$ min to eliminate thermal gradients at the test temperatures. Then cooled to test temperature where the specimen is held $1$ minute before deformation. After compression the specimen is quencted rapidly for microstructure studies. The procedure of this test is illustrated in Figure \ref{fig:Heating}.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]
{newFigures/Gleeble-3}
\caption{Gleeble 3800 simulator system}
\label{fig:Gleeble3800}
\end{figure}
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]
{newFigures/GleebleDesign}
\caption{Schematic diagram of the experimental process.}
\label{fig:Heating}
\end{figure}
\begin{table}[h!]
\centering{}
\caption{Chemical Composition of the AISI P20 Steel (Weight Percent)}
\scalebox{0.9}{
\begin{tabular}{lcccccccc}
\hline
C  &Mn & Mo   & Si   & Ni       &Cr& Cu & Other& Balance\\
\hline
 0.30        &       0.89      &0.52		   &	0.34     &	 0.68          &1.86& 0.17& Microalloying& Iron\\
\hline
\label{tab: Composition}
\end{tabular}}
\end{table}
%------------------------------------------------------------------------
\section{Results and discussion\label{sec:ConstLaws}}
%------------------------------------------------------------------------

%------------------------------------------------------------------------
\subsection{Flow stress behaviour}
%------------------------------------------------------------------------
The flow stress curves obtained from the compression tests performed on Gleeble-3800 simulator for each test condition are shown in Figure \ref{fig:rawData}. From these curves it can be seen that the flow stress increases with increasing strain rate but decreases with increasing temperature. It should be noted that the strain influences the flow stress. Indeed, for all strain values below $0.2$, the stress increases with strain before reaching a peak value at low strain rates and then decreases to maintain a stable value until the end of the test. The peak stress value increases with strain rate but decreases with temperature. For higher strain rates ($1.0\ \text{s}^{-1}$, $2.0\ \text{s}^{-1}$ and $5.0\ \text{s}^{-1}$) the flow stress increases throughout the test. The slight increase in stress at low strain rates when the strain is large is due to friction between the specimen and the grip during the test. For low strain rates, lubrication is lost over time and therefore friction occurs. This results in a progressive increase in stress.

To explain the overall behaviour of the flow stresses observed during this test, we will rely on two fundamental phenomena that can be observed during a compression test: Dynamic Recovery (DRC) and Dynamic Recrystallization (DRX). At the beginning of the deformation, the amount of dislocations increases and they start to move, leading to a rapid increase in stress known as work-hardening. As the strain continues to increase, the dislocations are set in motion, leading to a stress relaxation phenomenon usually known as DRC. During this DRC mechanism, some of the dislocations are removed, due to their combination with another dislocation, or due to their integration into a grain boundary, and some of them are rearranged and formed as sub-grains within the initial grains. The combination of these dislocations moderates the increase of dislocations linked to strain hardening and thus controls the amount of energy stored in the material and manages the increase in flow stress. DRX consists of the germination of new grains, at the borders or inside the initial grains. These two phenomena therefore explain the different shapes of the flow stress curves observed for this alloy. As an example, the Figure \ref{fig:BarNecking} , and Figure \ref{fig:AfterCompM} show the microstrutures of AISI P20 before and after compression. As observed on the flow curves, it can be seen that recrystallisation is almost complete for this temperature and strain rate. Indeed, most of the grain boundaries start to form new grains. The identification of the behavioural laws of these curves will be done in the next section in order to predict its behaviour in any test condition. This will be done on the one hand to test the ability of the models to describe an interpolation and on the other hand to do extrapolation.
{newFigures/rawData}
\caption{Flow stress-strain curves of AISI P20 alloy at various temperatures and strain rates.}
\label{fig:rawData}
{newFigures/BeforeCompM}
\caption{Optical micrograph of the AISI P20 steel before hot compression held during $5$ minutes at $1260^\circ$C}
\label{fig:BeforeCompM}
{newFigures/AfterCompM}
\caption{Optical micrograph of AISI P20 alloy after howt compression with a strain rate $0.001\ \text{s}^{-1}$ under $1150^\circ$C.}
\label{fig:AfterCompM}

@@

\subsection{Analytical constitutive flow laws}

@@

The complete characterization of the thermomechanical behavior of a material is very important before its use in numerical simulations. This characterization usually involves experimental tests that serve as a basis for the development of constitutive laws. As mentioned in the introduction, many constitutive laws exist today, and selecting one of them based on explicit criteria hinders the development of computational models for new materials. Since the purpose of this work is to compare analytical models with the ANN model in terms of interpolation and extrapolation, it is important to justify the choice of the analytical models presented here among the many that exist in the literature. Among the most widely used analytical models nowadays, the Johnson-Cook (JC) model is the most used due to its simplicity of use linked to the number of its reduced and easy to obtain parameters. It should be noted that it is not the best model but the most integrated in finite element software. Hence the existence of several modified forms of this model. The Zerilli-Armstrong model, and particularly its modified form (MZA), is also widely known since it tries to provide solutions to the shortcomings of the JC model, particularly with regard to the coupling of physical parameters. Hence its use also in this study. In this perspective, the Hansel Spittel model (HS)is also one of the most widely used models in numerical simulation of casting and forging operations as it is integrated in the FORGE software. Also, the method of identifying its parameters is very simple to do as it is done in one step. Of course, it has its limitations as will be shown in the following sections. Furthermore, the Arrhenius model will be used in this study because it is one of the models that not only provides a good prediction of the flow stress but also takes into account the physical parameter Q (the deformation activation energy coefficient) which is of great importance in the study related to the microstructural analysis of materials. Hence its use in the development of analytical models dedicated to the microstructural analysis of materials. The shortcomings of these models were the main motivation for the formulation of the model proposed in this work, the details of which can be found in our previous work.

@@

\subsubsection{Johnson--Cook model\label{sec:JC}}

@@

The JC model as mentioned above, is one of the most widely used analytical models as it can be applied to several materials under different conditions of strain, strain rate and temperature. However, the formulation of this model does not take into account the simultaneous effect of strain, strain rate and temperature. Indeed, it is formulated by describing the effect of each physical parameter (strain, strain rate and temperature) separately as a factor in the mathematical expression of the model. Hence its inability to describe the phenomenon of temperature-induced softening. The equation that describes this model is given as follows:

@@

From this equation one can easily obtain the parameter $C$ using the linear fitting method for each strain value and averaging all these values as shown in Figure \ref{fig:JCSigmaAB1} the value is $0.0976681$. 

@@

{newFigures/iCorrelationJC}
\caption{Comparison between the experimental and predicted flow stresses by JC model for interpolation estimation}
\label{fig:iCorrelationJC}

@@

{newFigures/eCorrelationJC}
\caption{Comparison between the experimental and predicted flow stresses by JC model for extrapolation estimation}
\label{fig:eCorrelationJC}
\caption{Parameters' constants of Johnson-Cook Model}
\scalebox{1.2}{
\begin{tabular}{lccccc}
&         &             &		   &		  &\\
&$A(\text{MPa})$    &$B(\text{MPa})$        & $n$     & $C$   & $m$       \\
&         &             &		   &	     &	    \\
Interpolation&$43.0988$&$114.518$& $0.313969$&$0.0993506$&$0.898932$\\
%\hdashline
Extrapolation&$43.0988$&$114.518$& $0.313969$&$0.0992144$&$0.910297$\\
\label{tab: JCparameters}
\subsubsection{The Modified Zerilli--Armstrong model\label{sec:MZA}}

>>>>

\title{Compararison of analytical and artificial neural network models in predicting hot deformation behavior of AISI P20 and  interpolation--extrapolation capacity validation}

\author[LGP]{Pierre Tize Mha}
\author[ETS]{Prashant Dhondapure}
\author[LGP]{Amèvi Tongne}
\author[ETS]{Mohammad Jahazi}
\author[LGP]{Olivier Pantalé \corref{cor1}}
\ead{Olivier.Pantale@enit.fr}
\ead[url]{http://www.enit.fr}

@@

\address[LGP]{Laboratoire Génie de Production, INP/ENIT, Université de Toulouse, 47 Av d'Azereix, Tarbes, France 65016}
\address[ETS]{Écolde de Technologie Supérieure, 1100 Rue Notre Dame O, Montréal, QC H3C 1K3}

\begin{abstract}
The high-temperature deformation behavior of the AISI P20 alloy was investigated by performing hot compression tests over a wide range of strains ($0.025 - 0.7$),  strain rates ($0.001\ \text{s}^{-1} - 5\ \text{s}^{-1}$ ) and temperatures ($1050^\circ\text{C} -1250^\circ$C ). Based on the experimental data (stress-strain curves), empirical models as the Johnson-Cook (JC), Modified-Zerilli-Armstrong (MZA), Hansel-Spittel (HS), Arhenius (AR) and PTM on the one hand and an ANN model on the other hand were developed to predict the flow behavior of this alloy. Then, a comparative study between the analytical models and the ANN model was carried out on the basis of the calculation of some coefficients allowing to test the efficiency of each model. These coefficients are the correlation coefficient $R$, the Average Absolute  Relative Error $AARE$ and the Root Mean Square Error ($RMSE$). From these parameters it was found that the JC, MZA and HS models are not suitable for predicting the behavior of AISI P20 alloy whereas the AR and PTM models can be used for most strain rates to predict the flow behavior of this alloy. For the strain rate ($\mdot{\varepsilon}^p = 0.01\ \text{s}^{-1}$)  the AR model fails to describe the flow for temperatures $1050^\circ$C and $1100^\circ$C  while in the case of the PTM model, the strain rate  ($\mdot{\varepsilon}^p = 0.1\ \text{s}^{-1}$) is not well described. This is due to the fact that this model is a sequence of polynomials which can create oscillations if the degree is high. Regarding the ANN model, the architecture retained is composed of two hidden layers of which the first hidden layer contains $15$ neurons and the second contains $7$ neurons and the input layer contains $3$ neurons corresponding to the strain ($\varepsilon^p$), strain rate ($\mdot{\varepsilon}^p$) and temperature ($T$) and the output layer contains a single neuron corresponding to the flow stress ($\sigma^y$). The results of the ANN model show a very good agreement between the predictions and the experiment. This demonstrates that this  formulation of constitutive law exceeds highly the analytical models in terms of prediction. To validate the performance of these models their capacities to interpolate and extrapolate the data were also tested where all the models relatively prove a good interpolation capacity whereas this is not the case for extrapolation where only the AR, PTM and ANN models are acceptable but still mentioning that the ANN model remains the best in both interpolation and extrapolation.  
\end{abstract}

\begin{keyword}
Zerilli--Armstrong flow law \sep Constitutive behavior \sep Arhenius--flow law \sep Hansel--Spittel flow law \sep Johnson--Cook flow law \sep interpolation \sep extrapolation \sep Artificial neural network
\end{keyword}

@@

%--------------------------------------------------------------------------
\section{Introduction\label{sec:Introduction}}
%--------------------------------------------------------------------------
Continuous casting and rolling mechanisms are an integral part of the technological methods of medium deformation through a continuous water-cooled crystalliser before entering a rolling mill. Some of the advantages of this method are increased metal yield, energy saving, improved quality of continuous casting and easier mechanisation and automation . Deformation being a very complex process involving both elastic and plastic deformation, is accompanied by high temperatures, high strain rates and complex and variable friction conditions \cite{he2013comparative}. However, a quantitative study and analysis of the deformation mechanism does not provide accurate analytical solutions . Moreover, conducting an experiment is not only time-consuming but also requires a high level of funding \cite{changizian2012high}. To mitigate these two factors (to name but a few), the Finite Element Method (FEM) is widely used in the analysis and optimisation of process parameters of various materials . FEM effectively compensates for the shortcomings of the analytical and experimental methods and plays an important role in the correct formulation of technological parameters, the improvement of product quality and the in-depth knowledge of the deformation mechanism. For example FEM has been widely used to simulate and analyse steel hot forming processes, such as forging, rolling, extrusion and casting but the accuracy of the numerical solution provided by FEM for this type of analysis strongly depends  on the accuracy of the constitutive equation of the concerned material \cite{qin2010modeling, mandal2009constitutive, ji2018constitutive}.

Indeed, hot forming processes are an essential step in the manufacturing of mechanical components and directly affect the ﬁnal quality of the products such as microstructure and mechanical properties \cite{ashtiani2012constitutive}. Several metallurgical phenomena such as work hardening and dynamic softening (dynamic recovery and recrystallisation), which very often occur during hot working, influence the quality of the deformation of materials. The complexity of the deformation behavior, however, leads to the fact that the deformation can only be calculated accurately by using a computer code to model the response of the material under the specified loading conditions. To this end, a mathematical formulation is often used to represent the flow behavior of materials in a form that can be used in finite element software \cite{lin2008prediction,wu2012flow}.

Based on a thorough search of the literature many researchers have tried to predict the mechanical behavior of materials under hot working conditions using constitutive models. Therefore, various constitutive equations have been proposed to describe the flow behavior including phenomenological, physical and artiﬁcial neural network (ANN) models \cite{lin2011critical,shin2010phenomenological,rusinek2010thermo}. As for the semi-empirical models, they try to be at the border between the empirical and physical models. Among a large number of models, the Johnson--Cook (JC) \cite{Johnson-1983-Constitutive, Johnson-1985-Fracture} and the Zerilli--Armstrong (ZA) \cite{Zerilli-1987-Dislocation} models are the best known and available in finite element codes. The JC model is widely used because it is simple to use and has fewer constants to be determined \cite{Khan-2004-Quasi-static, NematNasser-2003-Thermomechanical}. However, this model, in its original state, contains some problems related to the phenomena describing the behavior of the material. We refer here to strain hardening, strain rate hardening, and thermal effects that are described separately in the JC model as analyzed by Jia \eal \cite{Jia-2021-Modified-JC}. To overcome this difficulty, several modified forms of the Johnson--Cook model have been developed in the past \cite{Rule-1998-Revised-JC, Vural-2003-Large, Lin-2010-Modified-JC, Lin-2012-Phenomenological, Li-2013-Modified-JC, Zhou-2019-Research, Zhang-2015-Modified-JC}. According to its more physical formulation, the ZA model is preferred over the JC model because it allows for the coupling between strain rate and temperature effects \cite{Johnson-1988-Evaluation, Voyiadjis-2005-Microstructural, Dey-2007-Influence}. However, the drawbacks of this model are related to the temperature limit value of the test. Indeed, this model applies only to low strain rates and for temperatures limited to $0.6~T_m$ (where $T_m$ is the melting temperature of the material) \cite{Chiou-2005-Strain, Lee-2005-Strain, Chen-2007-Comparative, Lee-2006-Effects}. Moreover, this model, in its original form, does not take into account the coupling between deformation and temperature effects \cite{Samantaray-2009-Thermo-viscoplastic}. In order to reduce these shortcomings (\eg applying this model over a wider temperature range) of the ZA model, Modified Zerilli--Armstrong forms (MZA) have been proposed by several authors over the last past years \cite{Nemat-2004-Plasticity, Lennon-2004-Influence, Muralli-2017-Performance, Cheng-2021-Modified-ZA, Muralli-2021-Finite}.
The hyperbolic sine Arrhenius-type constitutive model is one of the most widely used phenomenological models both for the prediction of the hot deformation behavior of materials and for the analysis of the microstructure of the studied material because   this model allows to estimate the deformation energy of the material \cite{jonas1969strength,zhang2012effect,mostafaei2012hot}. Its original formulation has been revised several times in order to properly describe the obtained experimental behavior. In order to increase the accuracy of the model, the effects of strain \cite{slooff2007constitutive, li2012flow, xu2013constitutive}, strain rate \cite{lin2008constitutive, mandal2009constitutive} have been considered by several researchers. For example, Sloof \eal \cite{slooff2007constitutive} introduced a strain-dependent parameter into the initial formulation of the Arhenius model so that their model could predict the high temperature flow behavior of a magnesium alloy. Lin \eal \cite{lin2008constitutive}, on the other hand, modified the model by introducing a strain rate compensation factor.

The method based on artiﬁcial neural networks is an approach to predict the deformation behavior of materials under different conditions. ANN is in fact a method originally developed to simulate the biological processes of the human brain. And therefore it is not necessary to formulate a mathematical expresson and identify these parameters for this kind of model. The response of the deformation behavior of materials is highly non-linear at high temperatures, and many factors affecting the deformation behavior of materials are also non-linear; this makes the accuracy of the flow stress predicted by regression methods and the appropriate range is limited. Faced with these limitations, ANN models are therefore of interest.  Indeed, ANNs are particularly suited to deal with complex and non-linear relationships; consequently, ANNs have been successfully applied to predict the deformation behavior of materials under hot working conditions. ANNs can provide a fundamentally different approach to statistical or numerical techniques for materials modelling and materials processing control.

In this study, the high temperature deformation behavior of AISI P20 modified carbon steel subjected to compression tests will be highlighted. From the experimental data obtained, analytical models will be identified on the one hand in order to select the one that could better describe the behavior of this alloy and on the other hand, an ANN model is also developed and a comparative study between the analytical models and the ANN model will follow. A fundamental issue that will also be addressed in this work is the ability to interpolate and extrapolate of the developed models. Indeed, how do models behave when it comes to interpolating or extrapolating data that the model is never aware of ? the answer to this question will be well analyzed in this study.

The main parts of this work are as follows: Section \ref{sec:Materials} presents the material et experimental procedure inverstigated in this study. Section \ref{sec:ConstLaws} is devoted to the presentation of the models developed and their comparative. Section \ref{sec:inExtrapolation} details the validation of interpolation and extrapolation capacity of each model developped. Conclusion and future works are detailed in Section \ref{sec:Conclusion}.

%----------------------------------------------
\section{ Materials and Experiments \label{sec:Materials}}
%----------------------------------------------
%----------------------------------------------
\subsection{Experimental procedure}
%----------------------------------------------
The material used in this study is the modified carbon alloy AISI P20 whose chemical composition is given in the Table \ref{tab:Composition}. Cylindrical specimens were machined with an initial diameter of $10$ mm and a height of $15$ mm. Hot compression tests were conducted on a Gleeble-3800 thermo-simulator (Figure \ref{fig:Gleeble3800}), at temperatures of $1050^\circ$C,  $1050^\circ$C,  $1100^\circ$C,  $1150^\circ$C,  $1200^\circ$C and  $1250^\circ$C with strain rates of $0.001\ \text{s}^{-1}$, $0.01\ \text{s}^{-1}$, $0.1\ \text{s}^{-1}$, $1.0\ \text{s}^{-1}$, $2.0\ \text{s}^{-1}$ and $5.0\ \text{s}^{-1}$. Thin tantalum sheets were used as a lubricant material at the contact surface of anvils and specimen to minimize the frictions during the tests. The samples were heated up to temperature $1260^\circ$C  and were held for $5$ min to eliminate thermal gradients. Then cooled to test temperature where the specimen is held $1$ minute before deformation. After compression the specimen is quencted rapidly for microstructure studies. The procedure of this test is illustrated in Figure \ref{fig:Heating} where $h_0$ and $R_0$ are the height and radius before compression; $h$, $R_M$, and $R_T$ are respectively the height, big radius and small radius of the specimen after compression.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]
{newFigures/Gleeble-3}
\caption{Gleeble 3800 simulator system}
\label{fig:Gleeble3800}
\end{figure}
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]
{newFigures/GleebleDesign}
\caption{Schematic diagram of the experimental process.}
\label{fig:Heating}
\end{figure}
\begin{table}[h!]
\centering{}
\caption{Chemical Composition of the AISI P20 Steel (Weight Percent)}
\scalebox{0.9}{
\begin{tabular}{lcccccccc}
\hline
C  &Mn & Mo   & Si   & Ni       &Cr& Cu & Other& Balance\\
\hline
 0.30        &       0.89      &0.52		   &	0.34     &	 0.68          &1.86& 0.17& Microalloying& Iron\\
\hline
\label{tab:Composition}
\end{tabular}}
\end{table}
%------------------------------------------------------------------------
\subsection{Flow stress behavior}
%------------------------------------------------------------------------
The flow stress curves obtained from the compression tests performed on Gleeble-3800 simulator for each test condition are shown in Figure \ref{fig:rawData}. From these curves it can be seen that the flow stress increases with increasing strain rate but decreases with increasing temperature. It should be noted that the strain influences the flow stress. Indeed, for all strain values below $0.2$, the stress increases with strain before reaching a peak value at low strain rates and then decreases to maintain a stable value until the end of the test. The peak stress value increases with strain rate but decreases with temperature. For higher strain rates ($1.0\ \text{s}^{-1}$, $2.0\ \text{s}^{-1}$ and $5.0\ \text{s}^{-1}$) the flow stress increases throughout the test. The slight increase in stress at low strain rates when the strain is large is due to friction between the specimen and the grip during the test. For low strain rates, lubrication is lost over time and therefore friction occurs. This results in a progressive increase in stress.

To explain the overall behavior of the flow stresses observed during this test, we will rely on two fundamental phenomena that can be observed during a compression test: Dynamic Recovery (DRC) and Dynamic Recrystallization (DRX). At the beginning of the deformation, the amount of dislocations increases and they start to move, leading to a rapid increase in stress known as work-hardening. As the strain continues to increase, the dislocations are set in motion, leading to a stress relaxation phenomenon usually known as DRC. During this DRC mechanism, some of the dislocations are removed, due to their combination with another dislocation, or due to their integration into a grain boundary, and some of them are rearranged and formed as sub-grains within the initial grains. The combination of these dislocations moderates the increase of dislocations linked to strain hardening and thus controls the amount of energy stored in the material and manages the increase in flow stress. DRX consists of the germination of new grains, at the borders or inside the initial grains. These two phenomena therefore explain the different shapes of the flow stress curves observed for this alloy. As an example, the Figure \ref{fig:BeforeCompM} and Figure \ref{fig:AfterCompM} show the microstrutures of AISI P20 before and after compression. As observed on the flow curves, it can be seen that recrystallisation is almost complete for this temperature and strain rate. Indeed, most of the grain boundaries start to form new grains. The identification of the behavioral laws of these curves will be done in the next section in order to predict its behavior in any test condition.
\begin{figure}[!ht]
\centering
\includegraphics[width=1.02\columnwidth]
{newFigures/rawData}
\caption{Flow stress-strain curves of AISI P20 alloy at various temperatures and strain rates.}
\label{fig:rawData}
\end{figure}
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]
{newFigures/BeforeCompM}
\caption{Optical micrograph of the AISI P20 steel before hot compression held during $5$ minutes at $1260^\circ$C}
\label{fig:BeforeCompM}
\end{figure}
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]
{newFigures/AfterCompM}
\caption{Optical micrograph of AISI P20 alloy after howt compression with a strain rate $0.001\ \text{s}^{-1}$ under $1150^\circ$C.}
\label{fig:AfterCompM}
\end{figure}

@@

\section{Phenomenological constitutive flow laws\label{sec:ConstLaws}}

@@

The complete characterization of the thermomechanical behavior of a material is very important before its use in numerical simulations. This characterization usually involves experimental tests that serve as a basis for the development of constitutive laws. As mentioned in the introduction, many constitutive laws exist today, and selecting one of them based on explicit criteria hinders the development of computational models for new materials. Since the purpose of this work is to compare ANN model to analytical models it is important to justify the choice of the analytical models presented here among the many that exist in the literature. Among the most widely used analytical models nowadays, the Johnson-Cook (JC) model is the most used due to its simplicity of use linked to the number of its reduced and easy to obtain parameters. It should be noted that it is not the best model but the most integrated in finite element software. Hence the existence of several modified forms of this model. The Zerilli-Armstrong model, and particularly its modified form (MZA), is also widely known since it tries to provide solutions to the shortcomings of the JC model, particularly with regard to the coupling of physical parameters. Hence its use also in this study. In this perspective, the Hansel Spittel model (HS)is also one of the most widely used models in numerical simulation of casting and forging operations as it is integrated in the FORGE software. Also, the method of identifying its parameters is very simple to do as it is done in one step. Of course, it has its limitations as will be shown in the following sections. Furthermore, the Arrhenius model will be used in this study because it is one of the models that not only provides a good prediction of the flow stress but also takes into account the physical parameter Q (the deformation activation energy coefficient) which is of great importance in the study related to the microstructural analysis of materials. The PTM model is also used in this study as it is a model whose formulation does not impose a number of parameters a priori and has proven its ability in our previous work and therefore we want to test it on this type of material as it is formulated to fit any material. 

@@

\subsection{Johnson--Cook model\label{sec:JC}}

@@

The JC model as mentioned above, is one of the most widely used analytical models as it can be applied to several materials under different conditions of strain, strain rate and temperature . However, the formulation of this model does not take into account the simultaneous effect of strain, strain rate and temperature. Indeed, it is formulated by describing the effect of each physical parameter (strain, strain rate and temperature) separately as a factor in the mathematical expression of the model. Hence its inability to describe the phenomenon of temperature-induced softening. The equation that describes this model is given as follows \cite{johnson1985fracture}:

@@

From this equation one can easily obtain the parameter $C$ using the linear fitting method for each strain value and averaging all these values as shown in Figure \ref{fig:JCSigmaAB1} the value is $0.0995438$. 

@@

{newFigures/CompExpHS}
\caption{Comparison between the experimental and predicted flow stresses by HS model}

@@
